---
title: "Homework 5"
author: "Helinda He"
date: "2022-11-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Question 1

### a)
Men: M ~ N(4313, 583)

Women: W ~ N(5261, 807)

### b)
Leo: (4948 - 4313) / 583 = 1.089

Mary: (5513 - 5261) / 807 = 0.312

### c)
From the z value, Mary have a better result than Leo because Mary has a smaller z value. 

### d)
```{r}
1 - pnorm(q = 4948, mean = 4313, sd = 583)
```

Leo finished faster than 13.80% of the participants

### d)
```{r}
1 - pnorm(q = 5513, mean = 5261, sd = 807)
```

Mary finished faster than 37.74% of the participants. 

### e)
Yes, the probability will change since the distribution is not normal. 

# Question 2

### a) 
According to the Z-Table, the z value of the student who scored in the 80th percentile is z = 0.84. The observation will be 159.6729

### b)
According to the Z-Table, the z value of the student who scored greater than 30% of the test taker is z = -0.52. The observation will be 147.36. 


# Question 3

### a)
```{r}
qnorm(0.05, mean = 4313, sd = 583)
```

The time for the 5% fastest participants in men group is 3354.05

### b)
```{r}
qnorm(0.90, mean = 5261, sd = 807)
```
 
The time for the 10% slowest participants in women group is 6295.212. 

# Question 4

### a)
```{r}
1 - pnorm(83, mean = 77, sd = 5)
```

The probability will be 11.51%. 

### b)
```{r}
qnorm(0.1, mean = 77, sd = 5)
```

The coldest temperature is 70.59224 Fahrenheit. 


# Question 5

### a)
```{r}
pnorm(0, mean = 14.7, sd = 33)
```

The probability of the portfolio being negative return will be 0.3280. 

### b)
```{r}
qnorm(0.85, mean = 14.7, sd = 33)
```

The 15% of returns will be 48.90% percent.

# Question 6

### a)
Yes, it can be used the binomial distribution because it have independent samples, the probability is constant, and there is only two outcome can exist. 

### b)
```{r}
dbinom(97, 100, .90)
```

The probability of exactly 97 out of 100 randomly sampled American adults had chickenpox
during childhood is 0.0059. 

### c)
```{r}
dbinom(3, 100, .10)
```

The probability of exactly 3 out of a new sample of 100 American adults have not had
chickenpox in their childhood is also 0.0059. 

### d)
```{r}
1 - dbinom(0, 10, .90)
```

The probability of at least 1 out of 10 randomly sampled American adults have had chickenpox is 1. 

### e)
```{r}
pbinom(3, 10, .10)
```

The probability of at most 3 out of 10 randomly sampled American adults have not had
chickenpox is 0.9872. 


# Question 7

### a)
120 * 0.9 = 108
Var(people out of 120 that have chickenpox in their childhood) = 108 * 0.1 = 10.8
SD = $\sqrt{10.8} = 3.29$

### b)
No, because 105 is less than one standard deviation. 

### c)
```{r}
pbinom(105, 120, .9)
```

The probability of the people out of 120 that have chickenpox in their childhood is less than 105 is 0.2182. 

# Question 8

### a)
```{r}
dbinom(2, 3, .25)
```

The probability of two children who have the disease is 0.1406. 

### b)
```{r}
dbinom(0, 3, .25)
```

The probability of none of the three children will have the disease is 0.4219

### c)
```{r}
1 - dbinom(0, 3, 0.25)
```

The probability of at lease one of the three children do not carry the disease is 0.5781. 

### d)
$P(X = x) = (1-p)^{x-1} * P = (1 - 0.25)^{x-1} * (0.25)$
$P(X = 3) = (1 - 0.25)^{3 -1} * (0.25) = (0.75)^2 * (0.25) = 0.1406$

The probability of the fist children with the disease is the third child is 0.1406. 


# Question 9
Mean = 0.75 * 500 = 375
SD = $\sqrt{0.75 * 500 * (1 - 0.75)} = 9.6825$

### a)
```{r}
pnorm(400, mean = 375, sd = 9.6825) - pnorm(360, mean = 375, sd = 9.6825)
```

The probability between 400 and 600 of the drivers have the seat belt on is 0.9344. 

### b)
```{r}
pnorm(400, mean = 375, sd = 9.6825)
```

The probability of fewer than 400 people wearing the seat belt is 0.9951. 

### c)
```{r}
pnorm(400, mean = 375, sd = 9.6825)
```

The probability of at most there are 400 people wearing the seat belt is 0.9951. 


# Question 10

### a)
```{r}
pnorm(191, mean = 191, sd = 22.4)
```

The probability of the diameter of trees that is above 191 inches is 0.5.

### b)
```{r}
pnorm(190, mean = 191, sd = 22.4) - pnorm(180, mean = 191, sd = 22.4)
```

The probability of the diameter of trees between 180 and 190 inches is 0.1705. 

### c)
```{r}
qnorm(0.25, mean = 191, sd = 22.4)
```

The first 25% of the trees is 175.8914. 

### d)
```{r}
qnorm(0.75, mean = 191, sd = 22.4)
```

The third 25% of trees is 206.1086. 

### e)
The IQR will be 206.1086 - 175.8914, which will come up to 30.2172. 