---
title: "Lab 3"
author: "Helinda He"
date: "2022-11-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openintro)
library(statsr)
library(broom)
```

```{r}
hfi
```

# Exercise 1
*Question*: What are the dimensions of the dataset? What does each row represent?
We have 1458 rows and 123 columns. Each row represents the crime statistics in a particular region in a certain year. 
```{r}
dim(hfi)
glimpse(hfi)
```

# Exercise 2
*Question* : The data set spans a lot of years, but we are only interested in data from year 2016. Filter the data hfi data frame for year 2016, select the six variables, and assign the result to a data frame named hfi_2016.

```{r}
hfi_2016 <- hfi %>%
  # filter to choose rows
  filter(year == 2016) %>%
  # select to choose columns
  select(pf_score, pf_expression_control, hf_score)
```

# Exercise 3
*Question* : What type of plot would you use to display the relationship between the personal freedom score, pf_score, and pf_expression_control? Plot this relationship using the variable pf_expression_control as the predictor. Does the relationship look linear? If you knew a country’s pf_expression_control, or its score out of 10, with 0 being the most, of political pressures and controls on media content, would you be comfortable using a linear model to predict the personal freedom score?

*Answer*: A scatter plot is appropriate because both 'pf_score' and 'pf_expression_control' are numerical variables. Yes, this looks like a linear relationship because it looks like the points are following a straight line. Yes, we would 

```{r}
ggplot(data = hfi_2016, aes(x = pf_expression_control, y = pf_score)) + geom_point()
```

If the relationship is linear, we can quantify how strong this relationship is using the correlation coefficient. 

```{r}
hfi_2016 %>%
  summarize(cor(pf_expression_control, pf_score))
```


# Exercise 4
*Question*: Looking at your plot from the previous exercise, describe the relationship between these two variables. Make sure to discuss the form, direction, and strength of the relationship as well as any unusual observations.

The relationship between 'pf_expression_control' and 'pf_score' is linear, positive, and strong. 

# Exercise 5
*Question*: Using plot_ss, choose a line that does a good job of minimizing the sum of squares. Run the function several times. What was the smallest sum of squares that you got? How does it compare to your neighbours?

```{r}
plot_ss(x = pf_expression_control, 
        y = pf_score, 
        data = hfi_2016)
```

The 'lm()' function lets you creat a linear model.
The formula is: 'lm(y ~ x, data = dataset_name)'

```{r}
m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
tidy(m1)
```

*Interpretation*:

- To interpret intercept: When **x** is 0, **y** is expected to be **intercept**. 
Here, out intercept means: When pf_expression_control is 0, pf_score is expected to be 4.28. 

- To interpret slope: For every one unite increase in **x**, **y** is expected to **increase/decrease** by **slope**. 
Here, our slope means: For every one unit increase in pf_expression_control, pf_score is expected to increase by 0.542. 


```{r}
glance(m1)
```

We can assess model fit using $R^2$. We can interpret this as: *R^2*\% of the variability in **y** that is explained by **x**. In our case 71.4\% of the variability in the pf_score is explained by pf_expression_control

# Exercise 6
*Question*: Fit a new model that uses pf_expression_control to predict hf_score, or the total human freedom score. Using the estimates from the R output, write the equation of the regression line. What does the slope tell us in the context of the relationship between human freedom and the amount of political pressure on media content?

```{r}
lm(hf_score ~ pf_expression_control, data = hfi_2016)
```

The equation of the line is:
y = 5.0534 + 0.3684x

In this context, the slope means: for every


# Exercise 7
*Question*: If someone saw the least squares regression line and not the actual data, how would they predict a country’s personal freedom school for one with a 3 rating for pf_expression_control? Is this an overestimate or an underestimate, and by how much? In other words, what is the residual for this prediction?

```{r}
ggplot(data = hfi_2016, aes(x = pf_expression_control, y = pf_score)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

```{r}
4.28 + 0.542 * 3
```

We except pf_score to be 5.906 when pf_expression_control is 3. 

```{r}
hfi_2016 %>%
  filter(pf_expression_control == 3)
```

The actual value of pf_score is 5.47 when when pf_expression_control is 3. 

Our estimate line is an overestimate because the predicted value is 5.906, which is larger than the actual value of 5.47. 

The residual = Actual - Predicted

#Exercise 8#
*Question*: Is there any apparent pattern in the residuals plot? What does this indicate about the linearity of the relationship between the two variables?

```{r}
m1_aug <- augment(m1)
```

```{r}
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

It looks lie the residual-fitted plot shows a pattern. Specifically, the points are approaching the red line(0) from both directions as the fitted values increase. This indicates that the linearity condition may not be satisfied

#Exercise 9#
*Question*: Based on the histogram, does the nearly normal residuals condition appear to be violated? Why or why not?

```{r}
ggplot(data = m1_aug, aes(x = .resid)) + geom_histogram(binwidth = 0.25) + xlab("Residuals")
```

The distribution of the residuals may have a slight left skew, but overal loks relatively symmetric. The normality condition appears to be satisfied. 

#Exercise 10#
*Question*:Based on the residuals vs. fitted plot, does the constant variability condition appear to be violated? Why or why not?

The constant variance condition may not be satisfied because there seems to be a patter in the residual vs. fitted plot. 